### 2015-05-03 techwriting/CAI/reacha.tex
copy original prelim & theory

\section{Reachability Analysis of Finite State Machines (FSMs) using Algebraic Geometry}
% \subsection{An example illustrating our proposed approach}
In our approach we use symbolic state reachability with algebraic
geometry concepts. It is an abstraction based on word operand
definition of datapaths in circuits, and it can be applied
to arbitrary FSMs by bundling a set of bit-level variables together as
one or several word-level variables.  The abstraction polynomial,
encoding the reachable state space of the FSM, is obtained through
computing a GB of an elimination ideal using elimination term order
based on Theorem \ref{thm:elim}.

@@

\begin{Example}
For the circuit in Fig. \ref{fig:fsm} (b), we have the transition
functions of the machine as:

@@

In the next iteration, the algorithm uses state $S_1 = \{01\}$ as the
current (initial) state, and computes $S_2 = \{10\} = t_0\cdot
\overline{t_1}$ as the next reachable state, and so on. 
\end{Example}

@@

\begin{Definition}
\label{def:sum}
({\bf Sum/Product of Ideals}) If $I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$ are 
ideals in $\mathbb F[x_1, \dots, x_n]$, then the {\bf sum} of $I$ and $J$ is defined as
$$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$$ And the {\bf product} of $I$ and $J$ is defined
as
\begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle \nonumber
  \end{equation}
\end{Definition}

With concepts of ideal sums and products, we can obtain the
intersection and union of affine varieties as:
\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then ${\bf
  V}(I + J) = {\bf V}(I) \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) =
{\bf V}(I) \bigcup {\bf V}(J)$. 
\end{Theorem}

@@

\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb
  F[x_1, \dots, x_n]$, then $I:J$ is the set
  \begin{equation}
  \{f \in \mathbb F[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}\nonumber
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}

% However, complement of variety cannot be easily dealt by simple arithmetic on polynomial generators.
% For non-trivial cases we can only prove that 
% \begin{Theorem}
% Let $I, J$ be ideals in $\mathbb F[x_1,\dots,x_n]$, then 
% $${\bf V}(I:J) \supset {\bf V}({\bf I}({\bf V}(I) - {\bf V}(J)))$$
% \end{Theorem}
% Fortunately for most hardware verification cases, the form of polynomial generators are restricted.
% A proposition has been proved by \cite{jinpeng} that after adding {\bf vanishing polynomials ideal}
% the new composed ideal implies following corollary:

@@

\begin{Theorem}
\label{thm:quotient}
Let $I, J$ be ideals with vanishing polynomials over $\Fkk[x_1,\dots,x_n]$, then 
$${\bf V}(I:J) = {\bf V}(I) - {\bf V}(J)$$
\end{Theorem}

@@

\begin{Corollary}
$$V(J') = \overline{{\bf V}(J)} = {\bf V}(J_0:J)$$

\end{Corollary}

@@

\begin{Example}
We apply Algorithm \ref{alg:univa} to Ex.\ref{ex:motiv} to execute the FSM traversal. Let the initial state
$from^0 = S$ (i.e. state $\{00\}$), in the first iteration we compose an elimination ideal $J$ with

@@

% 
% $J = \langle t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1), t_1 - (xs_0+x+s_0s_1+s_0), S - s_0 - s_1, T - t_0 - t_1, S \rangle$
% and a vanishing polynomial ideal $J_0 = \langle x^2-x, s_0^2-s_0, s_1^2-s_1, t_0^2-t_0, t_1^2-t_1, S^4-S, T^4-T\rangle$
% The elimination term order is $\{x,s_0,s_1,t_0,t_1\} > S > T$. Compute a reduced GB for $J+J_0$, the result has 
% a polynomial generator with only $T$, we assign it to next state $to^1 = T^2+(\alpha+1)T+\alpha$ denoting
% $\{01,10\}$. Since formerly reached state $reach = T$, its complement is $\langle T^4-T\rangle:\langle T\rangle
% = T^3+1$ (i.e. states $\{01,10,11\}$). Then the newly reached state in this iteration is
% $\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle$. We add these states
% to formerly reached states $reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$
% (i.e. states $\{00,01,10\}$). We update the present states for next iteration $from^1 = S^2+(\alpha+1)S+\alpha$.
% In the second iteration, we just replace 1 generator in $J+J_0$: $S$ with $S^2+(\alpha+1)S+\alpha$, and elimination
% term order keeps the same. The reduced GB includes a polynomial generator $to^2 = T^2+\alpha T$ denotes states
% $\{00,10\}$. The complement of $reached$ is $\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
% = T + 1+\alpha$ (i.e. states $\{11\}$). We compute the newly reached state 
% $\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$, then the algorithm terminates and returns
% $reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$ as final reachable states.
\end{Example}

@@

\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=7.5in]{./new_multi.eps}
% \vspace{-0.2in}
\caption{A 3-bit RH-SMPO and its Moore FSM model}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}

@@

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
\DecMargin{1em}
\vspace{-0.1in}
\begin{Example}
\label{ex:RHSMPO}

@@

% By continue executing the loop,
Finally, after 3 iterations we obtain: $to_3 = \{ \mathbf{R'+A_{init}B_{init},}
~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
as the image. The final result is $from_3(R_{final}) = R_{final}+A_{init}\cdot
B_{init}$, which verifies the multiplier. 
\end{Example}
Using our approach we successfully verified the function of an 100-bit RH-multiplier,
which implies its capability to handle the verification of large designs. The results
are published in \cite{myDATE}.

>>>>

\section{Reachability Analysis of Finite State Machines (FSMs) using Algebraic Geometry}
% \subsection{An example illustrating our proposed approach}
In our approach we use symbolic state reachability with algebraic
geometry concepts. It is an abstraction based on word operand
definition of datapaths in circuits, and it can be applied
to arbitrary FSMs by bundling a set of bit-level variables together as
one or several word-level variables.  The abstraction polynomial,
encoding the reachable state space of the FSM, is obtained through
computing a GB of an elimination ideal using elimination term order
based on theorem \ref{thm:elim}.

@@

\begin{example}
For the circuit in Fig. \ref{fig:fsm} (b), we have the transition
functions of the machine as:

@@

In the next iteration, the algorithm uses state $S_1 = \{01\}$ as the
current (initial) state, and computes $S_2 = \{10\} = t_0\cdot
\overline{t_1}$ as the next reachable state, and so on. 
\end{example}

@@

\begin{definition}
\label{def:sum}
({\bf Sum/Product of Ideals}) If $I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$ are 
ideals in $\mathbb F[x_1, \dots, x_n]$, then the {\bf sum} of $I$ and $J$ is defined as
$$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$$ And the {\bf product} of $I$ and $J$ is defined
as
\begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle \nonumber
  \end{equation}
\end{definition}

With concepts of ideal sums and products, we can obtain the
intersection and union of affine varieties as:
\begin{theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then ${\bf
  V}(I + J) = {\bf V}(I) \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) =
{\bf V}(I) \bigcup {\bf V}(J)$. 
\end{theorem}

@@

\begin{definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb
  F[x_1, \dots, x_n]$, then $I:J$ is the set
  \begin{equation}
  \{f \in \mathbb F[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}\nonumber
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{definition}

% However, complement of variety cannot be easily dealt by simple arithmetic on polynomial generators.
% For non-trivial cases we can only prove that 
% \begin{theorem}
% Let $I, J$ be ideals in $\mathbb F[x_1,\dots,x_n]$, then 
% $${\bf V}(I:J) \supset {\bf V}({\bf I}({\bf V}(I) - {\bf V}(J)))$$
% \end{theorem}
% Fortunately for most hardware verification cases, the form of polynomial generators are restricted.
% A proposition has been proved by \cite{jinpeng} that after adding {\bf vanishing polynomials ideal}
% the new composed ideal implies following corollary:

@@

\begin{theorem}
\label{thm:quotient}
Let $I, J$ be ideals with vanishing polynomials over $\Fkk[x_1,\dots,x_n]$, then 
$${\bf V}(I:J) = {\bf V}(I) - {\bf V}(J)$$
\end{theorem}

@@

\begin{corollary}
$$V(J') = \overline{{\bf V}(J)} = {\bf V}(J_0:J)$$

\end{corollary}

@@

\begin{example}
We apply Algorithm \ref{alg:univa} to Ex.\ref{ex:motiv} to execute the FSM traversal. Let the initial state
$from^0 = S$ (i.e. state $\{00\}$), in the first iteration we compose an elimination ideal $J$ with

@@

% 
% $J = \langle t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1), t_1 - (xs_0+x+s_0s_1+s_0), S - s_0 - s_1, T - t_0 - t_1, S \rangle$
% and a vanishing polynomial ideal $J_0 = \langle x^2-x, s_0^2-s_0, s_1^2-s_1, t_0^2-t_0, t_1^2-t_1, S^4-S, T^4-T\rangle$
% The elimination term order is $\{x,s_0,s_1,t_0,t_1\} > S > T$. Compute a reduced GB for $J+J_0$, the result has 
% a polynomial generator with only $T$, we assign it to next state $to^1 = T^2+(\alpha+1)T+\alpha$ denoting
% $\{01,10\}$. Since formerly reached state $reach = T$, its complement is $\langle T^4-T\rangle:\langle T\rangle
% = T^3+1$ (i.e. states $\{01,10,11\}$). Then the newly reached state in this iteration is
% $\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle$. We add these states
% to formerly reached states $reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$
% (i.e. states $\{00,01,10\}$). We update the present states for next iteration $from^1 = S^2+(\alpha+1)S+\alpha$.
% In the second iteration, we just replace 1 generator in $J+J_0$: $S$ with $S^2+(\alpha+1)S+\alpha$, and elimination
% term order keeps the same. The reduced GB includes a polynomial generator $to^2 = T^2+\alpha T$ denotes states
% $\{00,10\}$. The complement of $reached$ is $\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
% = T + 1+\alpha$ (i.e. states $\{11\}$). We compute the newly reached state 
% $\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$, then the algorithm terminates and returns
% $reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$ as final reachable states.
\end{example}

@@

\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=4.5in]{./new_multi.eps}
% \vspace{-0.2in}
\caption{A 3-bit RH-SMPO and its Moore FSM model}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}

@@

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
\DecMargin{1em}
\vspace{-0.1in}
\begin{example}
\label{ex:RHSMPO}

@@

% By continue executing the loop,
Finally, after 3 iterations we obtain: $to_3 = \{ \mathbf{R'+A_{init}B_{init},}
~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
as the image. The final result is $from_3(R_{final}) = R_{final}+A_{init}\cdot
B_{init}$, which verifies the multiplier. 
\end{example}
Using our approach we successfully verified the function of an 100-bit RH-multiplier,
which implies its capability to handle the verification of large designs. The results
are published in \cite{myDATE}.
